\bf Interactive Visualization of Hierarchically Structured Data

### 2017-07-13 treelapse.tex
Add links + update figures
(red search lines now appear above blue select lines)

We have implemented our visualizations in D3, but encapsulated in an R package,
called treelapse, to facilitate rapid turnover from data preparation and
modeling to interactive exploration, and vice versa. Our code is open-source,
and available at \url{https://github.com/krisrs1128/treelapse}. We hope this
package encourages data analysts to work at the border between data modeling and
visualization, and more generally empowers a wider audience to apply less widely
known, but powerful, visualization ideas.

@@

We next consider an application unrelated to the microbiome, but with relatively
clear hierarchical structure. Our data are downloaded from
Zillow\footnote{\url{http://www.zillow.com/research/data/}}, and give the Zillow
Home Value Indexes at the neighborhood level, across the country, computed
monthly between 1996 and 2016. In our display, we have taken the base 10 log of
these indexes. As our hierarchical structure, we use each neighborhood's
assignment to state, regional, county, and city levels. We represent each of
these coarser spatial categories using the average of all neighborhoods
contained in them. We have filtered down to the 890 neighborhoods in California;
rendering more neighborhoods while keeping all 246 timepoints causes the
interface to lag\footnote{See below for potential optimizations, however.}.

@@

Our next example is a study in bikesharing demand, included as an
example of analyzing collections of time series when there is no obvious
hierarchical structure a priori. The data are available at the UCI
Repository\footnote{\url{https://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset}}
and were originally collected by a Washington D.C.-based bikesharing
system for use in a Kaggle prediction competition. The data are hourly
measurements of bike demand, aggregated across all bikesharing stations,
over two years, along with supplemental weather data. In the
competition, participants were asked to predict the hourly demand on a
held-out test set. Here, we adopt a descriptive view instead, attempting
to characterize factors associated with variation in bikesharing demand.

@@

{\centering \includegraphics[width=350px]{figure/weekend}}

\caption{By adjusting the two brushes in Figure \ref{fig:working}, we see that
  unlike weekday demand, weekend demand is unimodal. The few weekday series with
  unimodal series seem to be associated with holidays. This is the case for New
  Years' Eve, which is currently hovered over in the tree. The ease of
  transitioning from Figure \ref{fig:working} to this display indicates the
  importances of brushing in interaction.}\label{fig:weekend}
\end{figure}

\begin{figure}
{\centering \includegraphics[width=350px]{figure/weekday_2011}}

\caption{Weekday demand appears larger in 2012 than 2011 -- compare with Figure 
  \ref{fig:weekday2012}. Here, a brushes are introduced over the tree to see the
  series associated with a particular split point. The red nodes are the results
  of searches over the two nodes that are children of this split point. The fact
  that the \texttt{yr>=0.5} selected line is larger than the \texttt{yr<0.5}
  line means that demand was larger in 2012. In combination with searching and
  treeboxes, it is possible to interpret more subtle split points in the decision
  tree.}
\label{fig:weekday2011}
\end{figure}

{\centering \includegraphics[width=350px]{figure/weekday_2012}

\caption{Weekday demand increased in 2012 -- compare with 
  Figure \ref{fig:weekday2011}. The search terms are the same as in that figure,
  but the subtree associated with the 2012 split point is highlighted, using the
  union of two boxes. Note that unlike timebox trees, which highlighted series
  lying through the intersection of brushes, treeboxes highlight series within the
  union of brushes.}\label{fig:weekday2012}
\end{figure}

@@

>>>>

We have implemented our visualizations in D3, but encapsulated in an R package,
called treelapse, to facilitate rapid turnover from data preparation and
modeling to interactive exploration, and vice versa. Our code is open-source,
and is linked in Section \ref{sec:supp_links} of the supplementary materials. We
hope this package encourages data analysts to work at the border between data
modeling and visualization, and more generally empowers a wider audience to
apply less widely known, but powerful, visualization ideas.

@@

We next consider an application unrelated to the microbiome, but with relatively
clear hierarchical structure. Our data are downloaded from Zillow and give the
Zillow Home Value Indexes at the neighborhood level, across the country,
computed monthly between 1996 and 2016. A link to the data source is provided in
Section \ref{sec:supp_links} of the supplementary materials. In our display, we
have taken the base 10 log of these indexes. As our hierarchical structure, we
use each neighborhood's assignment to state, regional, county, and city levels.
We represent each of these coarser spatial categories using the average of all
neighborhoods contained in them. We have filtered down to the 890 neighborhoods
in California; rendering more neighborhoods while keeping all 246 timepoints
causes a severe lag in the interface.

@@

Our next example is a study in bikesharing demand, included as an example of
analyzing collections of time series when there is no obvious hierarchical
structure a priori. The data are available at the UCI Repository and were
originally collected by a Washington D.C.-based bikesharing system for use in a
Kaggle prediction competition. A link is provided in Section
\ref{sec:supp_links} of the supplementary materials. The data are hourly
measurements of bike demand, aggregated across all bikesharing stations, over
two years, along with supplemental weather data. In the competition,
participants were asked to predict the hourly demand on a held-out test set.
Here, we adopt a descriptive view instead, attempting to characterize factors
associated with variation in bikesharing demand.

@@

\begin{figure}
  \centering
  \includegraphics[width=350px]{figure/weekend}
    \caption{By adjusting the two brushes in Figure \ref{fig:working}, we see that
      unlike weekday demand, weekend demand is unimodal. The few weekday series with
      unimodal series seem to be associated with holidays. This is the case for New
      Years' Eve, which is currently hovered over in the tree. The ease of
      transitioning from Figure \ref{fig:working} to this display indicates the
      importances of brushing in interaction.}\label{fig:weekend}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=350px]{figure/weekday_2011}
    \caption{Weekday demand appears larger in 2012 than 2011 -- compare with Figure 
      \ref{fig:weekday2012}. Here, a brushes are introduced over the tree to see the
      series associated with a particular split point. The red nodes are the results
      of searches over the two nodes that are children of this split point. The fact
      that the \texttt{yr>=0.5} selected line is larger than the \texttt{yr<0.5}
      line means that demand was larger in 2012. In combination with searching and
      treeboxes, it is possible to interpret more subtle split points in the decision
      tree.} \label{fig:weekday2011}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=350px]{figure/weekday_2012}
  \caption{Weekday demand increased in 2012 -- compare with 
    Figure \ref{fig:weekday2011}. The search terms are the same as in that figure,
    but the subtree associated with the 2012 split point is highlighted, using the
    union of two boxes. Note that unlike timebox trees, which highlighted series
    lying through the intersection of brushes, treeboxes highlight series within the
    union of brushes.}\label{fig:weekday2012}
\end{figure}

@@

\section{Supplemental Materials}

\subsection{Links}
\label{sec:supp_links}

\begin{itemize}
\item treelapse home page: \url{https://krisrs1128.github.io/treelapse/}
\item treelapse development page: \url{https://github.com/krisrs1128/treelapse}
\item Home prices data: \url{http://www.zillow.com/research/data/}
\item Bikesharing data: \url{https://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset}
\end{itemize}
